\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{hyperref}
\usepackage{enumitem}

\title{ArXiv Research Assistant \\ \large CSCI-UA 473 --- Spring 2026}
\author{Kevin Pei, Kevin Dong, Ian Tang, Xan Carey, Vivek Kondapalli}
\date{}

\begin{document}
\maketitle

\section{Motivation}

ArXiv ingests thousands of new papers a day across cs.LG, cs.AI, cs.CV, cs.CL, cs.NE, and stat.ML.
For a researcher trying to map a subfield, lexical search is a hammer when you need a microscope ---
the query ``models that reduce transformer memory'' will miss a paper titled
``efficient attention via low-rank projections'' even though that is exactly the right hit.

We built three things on top of a curated 10{,}000-paper corpus:

\begin{enumerate}[noitemsep]
    \item Semantic question answering with grounded $[n]$ citations.
    \item An interactive 3D topic map where spatial neighborhoods correspond to research themes.
    \item A live ``Query Constellation'' that drops your query into the map and lights up its
          $k$ closest papers across the entire cloud.
\end{enumerate}

\section{Algorithms}

Every algorithm in the retrieval, clustering, and graph layers is implemented from scratch in NumPy.
No \texttt{sklearn}, no \texttt{faiss}, no \texttt{scipy.spatial}.

\subsection*{From-scratch}

\paragraph{Cosine similarity \& nearest-neighbor search} (\texttt{src/retrieval.py}).
For a query vector $\mathbf{q}$ and corpus matrix $X \in \mathbb{R}^{N \times D}$:
\[
\mathrm{sim}(\mathbf{q}, X_i) \;=\; \frac{\mathbf{q} \cdot X_i}{\|\mathbf{q}\|\,\|X_i\|}
\]
Top-$k$ selection uses \texttt{np.argpartition} for an $\mathcal{O}(N)$ partial selection
followed by an $\mathcal{O}(k \log k)$ sort over just the $k$ winners --- strictly better than
a full $\mathcal{O}(N \log N)$ sort when $k \ll N$. Two-stage retrieval drills from abstracts
into chunks/captions for the LLM context.

\paragraph{$k$-Means clustering} (\texttt{src/clustering.py}).
Lloyd's iteration with $k$-means$++$ initialization, $n\_\text{init}$ multi-restart keeping
best inertia, empty-cluster reseed-to-farthest-point, and a \emph{spherical} variant that
re-normalizes centroids after every update --- the right default on the unit sphere where
L2-normalized sentence-transformer embeddings live.

\paragraph{$k$-NN graph} (\texttt{src/graph.py}).
A symmetric undirected graph over cosine similarity:
\[
(i, j) \in E \iff j \in \mathrm{top}_k(i) \;\lor\; i \in \mathrm{top}_k(j),
\qquad
w_{ij} = \max\bigl(\mathrm{sim}(i,j),\ \mathrm{sim}(j,i)\bigr).
\]
Self-loops excluded via \texttt{np.fill\_diagonal(sim, $-\infty$)} before \texttt{argpartition}.
$k = 8$, chosen so the layout stays connected without becoming a hairball.

\paragraph{External cluster validation.}
Adjusted Rand Index and Normalized Mutual Information against arXiv \texttt{primary\_category},
both implemented from scratch and used to pick $k = 26$ clusters. Silhouette is uninformative
on 768-d embeddings; NMI peaks at $k=26$ then plateaus.

\subsection*{Pre-trained components (intentionally not from-scratch)}

\begin{itemize}[noitemsep]
    \item \textbf{Sentence encoder:} \texttt{all-mpnet-base-v2} (768-d, L2-normalized).
    \item \textbf{Cross-encoder reranker:} \texttt{ms-marco-MiniLM-L-6-v2}, fired conditionally
          when bi-encoder scores are weak (top-1 $< 0.5$) or bunched (spread $< 0.15$).
          Bi-encoder owns recall; cross-encoder owns precision.
    \item \textbf{Cluster labeler:} local Qwen2.5-1.5B-Instruct, 3-shot chat prompt over
          the 5 papers nearest each centroid.
    \item \textbf{Answer synthesis:} Claude Haiku 4.5 via OpenRouter, with a system prompt
          that forces $[n]$ citations to land at end-of-sentence.
    \item \textbf{Layout:} UMAP 3D, precomputed offline. A from-scratch Fruchterman-Reingold
          reference lives in the notebook --- it just doesn't scale to 10k nodes interactively.
\end{itemize}

\section{Tech Stack}

\begin{itemize}[noitemsep]
    \item \textbf{Backend:} FastAPI $+$ uvicorn. NumPy for every retrieval/clustering/graph
          primitive. \texttt{sentence-transformers} for encoding. SQLite for paper metadata
          with an LRU cache on top.
    \item \textbf{Frontend:} React $+$ Vite $+$ a \emph{custom} Three.js renderer.
          10k nodes and $\sim 20$k edges collapse into $\sim 3$ draw calls via
          \texttt{InstancedMesh} $+$ \texttt{LineSegments}, holding 60 fps on integrated GPUs.
    \item \textbf{LLM routing:} \texttt{LLM\_PROVIDER} env var swaps between local Qwen
          and OpenRouter Haiku per request. Cluster labeling stays pinned to the local model.
    \item \textbf{Data pipeline:} arXiv API $\to$ \texttt{ar5iv} HTML scrape $\to$ chunker
          $\to$ encoder $\to$ \texttt{abstracts.npy}, \texttt{chunks.npy}, \texttt{captions.npy}.
          $\sim 7$ hours from scratch, or 2 minutes from a cached release snapshot.
\end{itemize}

\section{References}
\begin{itemize}[noitemsep]
\end{itemize}

\end{document}
